Generating Random Logic Programs Using Constraint Programming

### 2020-08-10 seminar_talk/talk.tex
A bit about symmetry breaking

\documentclass{beamer}
\usepackage{tikz}
\usepackage[style=authoryear]{biblatex}
\usepackage{empheq}

\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes}
\tikzstyle{every picture}+=[remember picture]

@@

@@

@@

% width: 4.2679148611112
% height: 3.5575922222223

@@

\begin{frame}{Inference}
  \begin{minipage}[t][0.5\textheight]{\textwidth}
    Let $a \oplus b = a + b - ab$.

@@

    \only<5->{
      $\Pr[\textcolor{predicate}{\mathtt{smokes}}(\textcolor{constant}{\mathit{timothy}})]
      =
      \Pr[\textcolor{predicate}{\mathtt{stress}}(\textcolor{constant}{\mathit{timothy}})]
      = \textcolor{probability}{0.2}$
    }
  \end{minipage}
  \vfill
  \begin{overprint}
    \begin{onlyenv}<1>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{predicate}{\mathtt{cancer}}(\textcolor{variable}{\mathsf{P}})
        &\ifff
        \textcolor{predicate}{\mathtt{cancer\_spont}}(\textcolor{variable}{\mathsf{P}}).
        \\
        \textcolor{predicate}{\mathtt{cancer}}(\textcolor{variable}{\mathsf{P}})
        &\ifff
        \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{P}}),
        \textcolor{predicate}{\mathtt{cancer\_smoke}}(\textcolor{variable}{\mathsf{P}}).
      \end{empheq}
    \end{onlyenv}
    \begin{onlyenv}<2>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{probability}{0.1} \prob
        &\textcolor{predicate}{\mathtt{cancer\_spont}}(\textcolor{variable}{\mathsf{P}})
        \ifff
        \textcolor{predicate}{\mathtt{person}}(\textcolor{variable}{\mathsf{P}}).
        \\
        \textcolor{probability}{0.3} \prob
        &\textcolor{predicate}{\mathtt{cancer\_smoke}}(\textcolor{variable}{\mathsf{P}})
        \ifff
        \textcolor{predicate}{\mathtt{person}}(\textcolor{variable}{\mathsf{P}}).
      \end{empheq}
    \end{onlyenv}
    \begin{onlyenv}<3>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}})
        &\ifff
        \textcolor{predicate}{\mathtt{stress}}(\textcolor{variable}{\mathsf{X}}).
        \\
        \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}})
        &\ifff
        \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{Y}}),
        \textcolor{predicate}{\mathtt{influences}}(\textcolor{variable}{\mathsf{Y}},
        \textcolor{variable}{\mathsf{X}}).
      \end{empheq}
    \end{onlyenv}
    \begin{onlyenv}<4>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{probability}{0.2} \prob &\textcolor{predicate}{\mathtt{stress}}(\textcolor{variable}{\mathsf{P}}) \ifff \textcolor{predicate}{\mathtt{person}}(\textcolor{variable}{\mathsf{P}}). \\
        \textcolor{probability}{0.3} \prob &\textcolor{predicate}{\mathtt{influences}}(\textcolor{variable}{\mathsf{P}_1}, \textcolor{variable}{\mathsf{P}_2}) \ifff \textcolor{predicate}{\mathtt{friend}}(\textcolor{variable}{\mathsf{P}_1}, \textcolor{variable}{\mathsf{P}_2}).
      \end{empheq}
    \end{onlyenv}
    \begin{onlyenv}<5>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{probability}{0.2} \prob &\textcolor{predicate}{\mathtt{stress}}(\textcolor{variable}{\mathsf{P}}) \ifff \textcolor{predicate}{\mathtt{person}}(\textcolor{variable}{\mathsf{P}}). \\
        &\textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}}) \ifff \textcolor{predicate}{\mathtt{stress}}(\textcolor{variable}{\mathsf{X}}).
      \end{empheq}
    \end{onlyenv}
  \end{overprint}
\end{frame}
% TODO: colours for numbers/probabilities

% TODO: what properties does d-DNNF ensure?
% TODO: what about SDD?
% TODO: explain how each of them works
\begin{frame}{Inference Algorithms and Knowledge Compilation Maps}
  \begin{description}
  \item[NNF] negation normal form
  \item[d-DNNF] deterministic decomposable negation normal form
  \item[BDD] binary decision diagrams
  \item[SDD] sentential decision diagrams
  \item[$k$-Best] only use the $k$ most probable proofs
  \end{description}
\end{frame}

\begin{frame}{Example Diagrams for $C \land (A \lor \neg B)$}
  \begin{columns}[b]
    \begin{column}{0.33\textwidth}
      \begin{figure}
        \centering
        \begin{tikzpicture}[level distance=0.8cm]
          \node[draw,circle] {$\land$}
          child {node[draw,rectangle] {$C$}}
          child {node[draw,circle] {$\lor$}
            child {node[draw,rectangle] {$A$}}
            child {node[draw,rectangle] {$\neg B$}
            }
          };
        \end{tikzpicture}
        \caption{NNF}
      \end{figure}
      \vspace{-1em}
      \begin{figure}
      \centering
        \begin{tikzpicture}[level distance=0.8cm]
          \node[draw,circle] (A) {$A$}
          child {edge from parent[draw=none]}
          child {node[draw,circle] (B) {$B$} edge from parent[dashed]
            child {node[draw,circle,solid] (C) {$C$} edge from parent[dashed]
              child {node[draw,rectangle,solid] {$1$} edge from parent[solid]}
              child {node[draw,rectangle,solid] (0) {$0$}}
            }
            child {edge from parent[draw=none]}
          };
          \draw (A) -- (C);
          \draw (B) -- (0);
        \end{tikzpicture}
        \caption{BDD}
      \end{figure}
    \end{column}
    \begin{column}{0.33\textwidth}
      \begin{figure}
        \centering
        \begin{tikzpicture}[level distance=0.8cm]
          \node[draw,circle] {$\land$}
          child {node[draw,circle] {$\lor$}
            child {node[draw,circle] (parent) {$\land$}
              child {node[draw,rectangle] {$B$}}
              child {edge from parent[draw=none]}
            }
            child {node[draw,circle] {$\land$}
              child {node[draw,circle] {$\lor$}
                child {node[draw,rectangle] (A) {$A$}}
                child {node[draw,rectangle] {$\neg A$}}
              }
              child {node[draw,rectangle] {$\neg B$}}
            }
          }
          child {node[draw,rectangle] {$C$}};
          \draw (parent) -- (A);
        \end{tikzpicture}
        \caption{d-DNNF}
      \end{figure}
    \end{column}
    \begin{column}{0.33\textwidth}
      \begin{figure}
        \centering
        \begin{tikzpicture}[level distance=0.8cm]
          \tikzset{
            mysplit/.style={
              draw,
              rectangle,
              rectangle split,
              rectangle split horizontal,
              rectangle split parts=2
            }
          }
          \node[draw,circle] {$1$}
          child {node[mysplit] (bullet) {
              \nodepart{one} $\neg A$
              \nodepart{two}
            }
            child {node[draw,circle] (3) {$3$} edge from parent[draw=none]
              child {node[mysplit] {
                  \nodepart{one} $\neg B$
                  \nodepart{two} $C$
                }
              }
              child {node[mysplit] {
                  \nodepart{one} $B$
                  \nodepart{two} $0$
                }
              }
            }
          }
          child {node[mysplit] {
              \nodepart{one} $A$
              \nodepart{two} $C$
            }};
          \draw[*-] let \p1 = (bullet.two), \p2 = (bullet.center) in ({\x1 + 2.5},{\y2 + 2}) -- (3);
        \end{tikzpicture}
        \caption{SDD}
      \end{figure}
      \vspace{-2em}
      \begin{figure}
      \centering
      \begin{tikzpicture}[level distance=0.8cm]
        \node[draw,circle] {$1$}
        child {node[draw,circle] {$0$}
          child {node[draw,rectangle] {$A$}}
        }
        child {node[draw,circle] {$3$}
          child {node[draw,circle] {$2$}
            child {node[draw,rectangle] {$B$}}
          }
          child {node[draw,circle] {$4$}
            child {node[draw,rectangle] {$C$}}
          }
        };
      \end{tikzpicture}
      \caption{vtree}
      \end{figure}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}{What Characterizes a (Probabilistic) Logic Program?}
  \begin{columns}
    \hspace*{-0.7cm}\begin{column}{0.75\textwidth}
      \begin{empheq}[left =\onslide<6->{\color{color5}\empheqlbrace}]{equation}
        \begin{align*}
      \textcolor<5->{probability}{0.2} \prob &\textcolor<2->{predicate}{\mathtt{stress}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
      \textcolor<5->{probability}{0.3} \prob &\textcolor<2->{predicate}{\mathtt{influences}}(\textcolor<3->{variable}{\mathsf{P}_1}, \textcolor<3->{variable}{\mathsf{P}_2}) \ifff \textcolor<2->{predicate}{\mathtt{friend}}(\textcolor<3->{variable}{\mathsf{P}_1}, \textcolor<3->{variable}{\mathsf{P}_2}). \\
      \textcolor<5->{probability}{0.1} \prob &\textcolor<2->{predicate}{\mathtt{cancer\_spont}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
      \textcolor<5->{probability}{0.3} \prob &\textcolor<2->{predicate}{\mathtt{cancer\_smoke}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{X}}) \ifff \textcolor<2->{predicate}{\mathtt{stress}}(\textcolor<3->{variable}{\mathsf{X}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{X}}) \ifff \textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{Y}}), \textcolor<2->{predicate}{\mathtt{influences}}(\textcolor<3->{variable}{\mathsf{Y}}, \textcolor<3->{variable}{\mathsf{X}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{cancer}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{cancer\_spont}}(\textcolor<3->{variable}{\mathsf{P}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{cancer}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \tikz \coordinate (start);\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{P}}), \textcolor<2->{predicate}{\mathtt{cancer\_smoke}}(\textcolor<3->{variable}{\mathsf{P}})\tikz \coordinate (end);. \\
                                         &\textcolor<2->{predicate}{\mathtt{person}}(\textcolor<4->{constant}{\mathit{michelle}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{person}}(\textcolor<4->{constant}{\mathit{timothy}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{friend}}(\textcolor<4->{constant}{\mathit{timothy}}, \textcolor<4->{constant}{\mathit{michelle}}).
        \end{align*}
      \end{empheq}
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{itemize}
      \item[\textcolor{predicate}{\textbullet}]<2-> predicates, arities
      \item[\textcolor{variable}{\textbullet}]<3-> variables
      \item[\textcolor{constant}{\textbullet}]<4-> constants
      \item[\textcolor{probability}{\textbullet}]<5-> probabilities
      \item[\textcolor{color5}{\textbullet}]<6-> length
      \item[\textcolor{color6}{\textbullet}]<7-> complexity
      \end{itemize}
    \end{column}
  \end{columns}
  \onslide<7->{
    \begin{tikzpicture}[overlay]
      \draw [decorate,decoration={brace,amplitude=10pt,mirror},color=color6] (start) -- (end);
    \end{tikzpicture}
  }
\end{frame}

\begin{frame}{Clauses As Trees}
  \[
    \neg \textcolor{predicate}{\mathtt{p}}(\textcolor{variable}{\mathsf{X}})
    \lor (\textcolor{predicate}{\mathtt{q}}(\textcolor{variable}{\mathsf{X}})
    \land \textcolor{predicate}{\mathtt{r}}(\textcolor{variable}{\mathsf{X}}))
  \]
  \vfill
  \begin{columns}
    \begin{column}{0.4\textwidth}
      \pause
      \centering
      \begin{tikzpicture}[level 1/.style={sibling distance=30mm},level 2/.style={sibling distance=15mm}]
        \node[draw,circle] {$\lor$}
        child {node[draw,circle] {$\neg$}
          child {node[draw,rectangle] {$\textcolor{predicate}{\mathtt{p}}(\textcolor{variable}{\mathsf{X}})$}}
        }
        child {node[draw,circle] {$\land$}
          child {node[draw,rectangle] {$\textcolor{predicate}{\mathtt{q}}(\textcolor{variable}{\mathsf{X}})$}}
          child {node[draw,rectangle] {$\textcolor{predicate}{\mathtt{r}}(\textcolor{variable}{\mathsf{X}})$}}
        };
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.6\textwidth}
      \pause
      \centering
      \begin{tabular}{|c|c|c|c|c|c|c|}
        \hline
        0 & 0 & 0 & 1 & 2 & 2 & 6 \\
        \hline
        $\lor$ & $\neg$ & $\land$ & $\textcolor{predicate}{\mathtt{p}}(\textcolor{variable}{\mathsf{X}})$ & $\textcolor{predicate}{\mathtt{q}}(\textcolor{variable}{\mathsf{X}})$ & $\textcolor{predicate}{\mathtt{r}}(\textcolor{variable}{\mathsf{X}})$ & $\top$ \\
        \hline
      \end{tabular}
    \end{column}
  \end{columns}
\end{frame}

>>>>

\documentclass{beamer}
\usepackage{tikz}
\usepackage[style=authoryear]{biblatex}
\usepackage{empheq}
\usepackage{pifont}

@@

\usetikzlibrary{arrows}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{positioning}
\usetikzlibrary{shapes}
\tikzstyle{every picture}+=[remember picture]
\tikzset{onslide/.code args={<#1>#2}{%
  \only<#1>{\pgfkeysalso{#2}}
}}

@@

\newcommand{\cmark}{\textcolor{green!50!black}{\ding{51}}}%
\newcommand{\xmark}{\textcolor{red}{\ding{55}}}%

@@

\begin{frame}
  \frametitle{Outline}
  \tableofcontents
\end{frame}

\section{Probabilistic Logic Programming}

@@

\begin{frame}{Probabilistic Inference}
  \begin{minipage}[t][0.5\textheight]{\textwidth}
    Let \alert{$a \oplus b = a + b - ab$}.

@@

    \only<5->{
      $\Pr[\textcolor{predicate}{\mathtt{smokes}}(\textcolor{constant}{\mathit{timothy}})]
      =
      \Pr[\textcolor{predicate}{\mathtt{stress}}(\textcolor{constant}{\mathit{timothy}})]
      = \textcolor{probability}{0.2}$
    }
  \end{minipage}
  \vfill
  \begin{overprint}
    \begin{onlyenv}<1>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{predicate}{\mathtt{cancer}}(\textcolor{variable}{\mathsf{P}})
        &\ifff
        \textcolor{predicate}{\mathtt{cancer\_spont}}(\textcolor{variable}{\mathsf{P}}).
        \\
        \textcolor{predicate}{\mathtt{cancer}}(\textcolor{variable}{\mathsf{P}})
        &\ifff
        \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{P}}),
        \textcolor{predicate}{\mathtt{cancer\_smoke}}(\textcolor{variable}{\mathsf{P}}).
      \end{empheq}
    \end{onlyenv}
    \begin{onlyenv}<2>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{probability}{0.1} \prob
        &\textcolor{predicate}{\mathtt{cancer\_spont}}(\textcolor{variable}{\mathsf{P}})
        \ifff
        \textcolor{predicate}{\mathtt{person}}(\textcolor{variable}{\mathsf{P}}).
        \\
        \textcolor{probability}{0.3} \prob
        &\textcolor{predicate}{\mathtt{cancer\_smoke}}(\textcolor{variable}{\mathsf{P}})
        \ifff
        \textcolor{predicate}{\mathtt{person}}(\textcolor{variable}{\mathsf{P}}).
      \end{empheq}
    \end{onlyenv}
    \begin{onlyenv}<3>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}})
        &\ifff
        \textcolor{predicate}{\mathtt{stress}}(\textcolor{variable}{\mathsf{X}}).
        \\
        \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}})
        &\ifff
        \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{Y}}),
        \textcolor{predicate}{\mathtt{influences}}(\textcolor{variable}{\mathsf{Y}},
        \textcolor{variable}{\mathsf{X}}).
      \end{empheq}
    \end{onlyenv}
    \begin{onlyenv}<4>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{probability}{0.2} \prob &\textcolor{predicate}{\mathtt{stress}}(\textcolor{variable}{\mathsf{P}}) \ifff \textcolor{predicate}{\mathtt{person}}(\textcolor{variable}{\mathsf{P}}). \\
        \textcolor{probability}{0.3} \prob &\textcolor{predicate}{\mathtt{influences}}(\textcolor{variable}{\mathsf{P}_1}, \textcolor{variable}{\mathsf{P}_2}) \ifff \textcolor{predicate}{\mathtt{friend}}(\textcolor{variable}{\mathsf{P}_1}, \textcolor{variable}{\mathsf{P}_2}).
      \end{empheq}
    \end{onlyenv}
    \begin{onlyenv}<5>
      \begin{empheq}[box=\fbox]{align*}
        \textcolor{probability}{0.2} \prob &\textcolor{predicate}{\mathtt{stress}}(\textcolor{variable}{\mathsf{P}}) \ifff \textcolor{predicate}{\mathtt{person}}(\textcolor{variable}{\mathsf{P}}). \\
        &\textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}}) \ifff \textcolor{predicate}{\mathtt{stress}}(\textcolor{variable}{\mathsf{X}}).
      \end{empheq}
    \end{onlyenv}
  \end{overprint}
\end{frame}

\begin{frame}{Inference Algorithms and Knowledge Compilation Maps}
  \begin{description}
  \item[NNF] negation normal form
  \item[BDD] binary decision diagrams
  \item[SDD] sentential decision diagrams
  \item[$k$-Best] only use the \alert{$k$} most probable proofs
  \item[d-DNNF] deterministic decomposable negation normal form
    \begin{itemize}
    \item<2-> for every pair \alert{$\alpha \lor \beta$}, we have \alert{$\alpha
        \land \beta = \bot$}
    \item<2-> for every pair \alert{$\alpha \land \beta$}, no atoms are shared
      between \alert{$\alpha$} and \alert{$\beta$}
    \item[\xmark\xmark]<3> $(A \lor C) \land (A \lor \neg B)$
    \item[\xmark\cmark]<3> $C \land (A \lor \neg B)$
    \item[\cmark\xmark]<3> $B \land C \land [(B \land A) \lor \neg B]$
    \item[\cmark\cmark]<3> $C \land [(B \land A) \lor \neg B]$
    \end{itemize}
  \end{description}
\end{frame}

\begin{frame}{Example Diagrams for \alert{$C \land (A \lor \neg B)$}}
  \begin{columns}[b]
    \begin{column}{0.33\textwidth}
      \begin{figure}
        \centering
        \begin{tikzpicture}[level distance=0.8cm]
          \node[draw,circle] {$\land$}
          child {node[draw,rectangle] {$C$}}
          child {node[draw,circle] {$\lor$}
            child {node[draw,rectangle] {$A$}}
            child {node[draw,rectangle] {$\neg B$}
            }
          };
        \end{tikzpicture}
        \caption{NNF}
      \end{figure}
      \vspace{-1em}
      \begin{figure}
      \centering
        \begin{tikzpicture}[level distance=0.8cm]
          \node[draw,circle] (A) {$A$}
          child {edge from parent[draw=none]}
          child {node[draw,circle] (B) {$B$} edge from parent[dashed]
            child {node[draw,circle,solid] (C) {$C$} edge from parent[dashed]
              child {node[draw,rectangle,solid] {$1$} edge from parent[solid]}
              child {node[draw,rectangle,solid] (0) {$0$}}
            }
            child {edge from parent[draw=none]}
          };
          \draw (A) -- (C);
          \draw (B) -- (0);
        \end{tikzpicture}
        \caption{BDD}
      \end{figure}
    \end{column}
    \begin{column}{0.33\textwidth}
      \begin{figure}
        \centering
        \begin{tikzpicture}[level distance=0.8cm]
          \node[draw,circle] {$\land$}
          child {node[draw,circle] {$\lor$}
            child {node[draw,circle] (parent) {$\land$}
              child {node[draw,rectangle] {$B$}}
              child {edge from parent[draw=none]}
            }
            child {node[draw,circle] {$\land$}
              child {node[draw,circle] {$\lor$}
                child {node[draw,rectangle] (A) {$A$}}
                child {node[draw,rectangle] {$\neg A$}}
              }
              child {node[draw,rectangle] {$\neg B$}}
            }
          }
          child {node[draw,rectangle] {$C$}};
          \draw (parent) -- (A);
        \end{tikzpicture}
        \caption{d-DNNF}
      \end{figure}
    \end{column}
    \begin{column}{0.33\textwidth}
      \begin{figure}
        \centering
        \begin{tikzpicture}[level distance=0.8cm]
          \tikzset{
            mysplit/.style={
              draw,
              rectangle,
              rectangle split,
              rectangle split horizontal,
              rectangle split parts=2
            }
          }
          \node[draw,circle] {$1$}
          child {node[mysplit] (bullet) {
              \nodepart{one} $\neg A$
              \nodepart{two}
            }
            child {node[draw,circle] (3) {$3$} edge from parent[draw=none]
              child {node[mysplit] {
                  \nodepart{one} $\neg B$
                  \nodepart{two} $C$
                }
              }
              child {node[mysplit] {
                  \nodepart{one} $B$
                  \nodepart{two} $\bot$
                }
              }
            }
          }
          child {node[mysplit] {
              \nodepart{one} $A$
              \nodepart{two} $C$
            }};
          \draw[*-] let \p1 = (bullet.two), \p2 = (bullet.center) in ({\x1 + 2.5},{\y2 + 2}) -- (3);
        \end{tikzpicture}
        \caption{SDD}
      \end{figure}
      \vspace{-2em}
      \begin{figure}
      \centering
      \begin{tikzpicture}[level distance=0.8cm]
        \node[draw,circle] {$1$}
        child {node[draw,circle] {$0$}
          child {node[draw,rectangle] {$A$}}
        }
        child {node[draw,circle] {$3$}
          child {node[draw,circle] {$2$}
            child {node[draw,rectangle] {$B$}}
          }
          child {node[draw,circle] {$4$}
            child {node[draw,rectangle] {$C$}}
          }
        };
      \end{tikzpicture}
      \caption{vtree}
      \end{figure}
    \end{column}
  \end{columns}
\end{frame}

\section{The Constraint Model}

\begin{frame}{What Characterises a (Probabilistic) Logic Program?}
  \begin{columns}
    \hspace*{-0.7cm}\begin{column}{0.75\textwidth}
      \begin{empheq}[left =\onslide<6->{\color{color5}\empheqlbrace}]{equation}
        \begin{align*}
      \textcolor<5->{probability}{0.2} \prob &\textcolor<2->{predicate}{\mathtt{stress}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
      \textcolor<5->{probability}{0.3} \prob &\textcolor<2->{predicate}{\mathtt{influences}}(\textcolor<3->{variable}{\mathsf{P}_1}, \textcolor<3->{variable}{\mathsf{P}_2}) \ifff \textcolor<2->{predicate}{\mathtt{friend}}(\textcolor<3->{variable}{\mathsf{P}_1}, \textcolor<3->{variable}{\mathsf{P}_2}). \\
      \textcolor<5->{probability}{0.1} \prob &\textcolor<2->{predicate}{\mathtt{cancer\_spont}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
      \textcolor<5->{probability}{0.3} \prob &\textcolor<2->{predicate}{\mathtt{cancer\_smoke}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{person}}(\textcolor<3->{variable}{\mathsf{P}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{X}}) \ifff \textcolor<2->{predicate}{\mathtt{stress}}(\textcolor<3->{variable}{\mathsf{X}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{X}}) \ifff \textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{Y}}), \textcolor<2->{predicate}{\mathtt{influences}}(\textcolor<3->{variable}{\mathsf{Y}}, \textcolor<3->{variable}{\mathsf{X}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{cancer}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \textcolor<2->{predicate}{\mathtt{cancer\_spont}}(\textcolor<3->{variable}{\mathsf{P}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{cancer}}(\textcolor<3->{variable}{\mathsf{P}}) \ifff \tikz \coordinate (start);\textcolor<2->{predicate}{\mathtt{smokes}}(\textcolor<3->{variable}{\mathsf{P}}), \textcolor<2->{predicate}{\mathtt{cancer\_smoke}}(\textcolor<3->{variable}{\mathsf{P}})\tikz \coordinate (end);. \\
                                         &\textcolor<2->{predicate}{\mathtt{person}}(\textcolor<4->{constant}{\mathit{michelle}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{person}}(\textcolor<4->{constant}{\mathit{timothy}}). \\
                                         &\textcolor<2->{predicate}{\mathtt{friend}}(\textcolor<4->{constant}{\mathit{timothy}}, \textcolor<4->{constant}{\mathit{michelle}}).
        \end{align*}
      \end{empheq}
    \end{column}
    \begin{column}{0.25\textwidth}
      \begin{itemize}
      \item[\textcolor{predicate}{\textbullet}]<2-> predicates, arities
      \item[\textcolor{variable}{\textbullet}]<3-> variables
      \item[\textcolor{constant}{\textbullet}]<4-> constants
      \item[\textcolor{probability}{\textbullet}]<5-> probabilities
      \item[\textcolor{color5}{\textbullet}]<6-> length
      \item[\textcolor{color6}{\textbullet}]<7-> complexity
      \end{itemize}
    \end{column}
  \end{columns}
  \onslide<7->{
    \begin{tikzpicture}[overlay]
      \draw [decorate,decoration={brace,amplitude=10pt,mirror},color=color6] (start) -- (end);
    \end{tikzpicture}
  }
\end{frame}

% TODO: I could talk about the constraints behind this
\begin{frame}{Clauses As Trees}
  \[
    \neg \textcolor{predicate}{\mathtt{p}}(\textcolor{variable}{\mathsf{X}})
    \lor (\textcolor{predicate}{\mathtt{q}}(\textcolor{variable}{\mathsf{X}})
    \land \textcolor{predicate}{\mathtt{r}}(\textcolor{variable}{\mathsf{X}}))
  \]
  \vfill
  \begin{columns}
    \begin{column}{0.4\textwidth}
      \pause
      \centering
      \begin{tikzpicture}[level 1/.style={sibling distance=30mm},level 2/.style={sibling distance=15mm}]
        \node[draw,circle] {$\lor$}
        child {node[draw,circle] {$\neg$}
          child {node[draw,rectangle] {$\textcolor{predicate}{\mathtt{p}}(\textcolor{variable}{\mathsf{X}})$}}
        }
        child {node[draw,circle] {$\land$}
          child {node[draw,rectangle] {$\textcolor{predicate}{\mathtt{q}}(\textcolor{variable}{\mathsf{X}})$}}
          child {node[draw,rectangle] {$\textcolor{predicate}{\mathtt{r}}(\textcolor{variable}{\mathsf{X}})$}}
        };
      \end{tikzpicture}
    \end{column}
    \begin{column}{0.6\textwidth}
      \pause
      \centering
      \begin{tabular}{|c|c|c|c|c|c|c|}
        \hline
        0 & 0 & 0 & 1 & 2 & 2 & 6 \\
        \hline
        $\lor$ & $\neg$ & $\land$ & $\textcolor{predicate}{\mathtt{p}}(\textcolor{variable}{\mathsf{X}})$ & $\textcolor{predicate}{\mathtt{q}}(\textcolor{variable}{\mathsf{X}})$ & $\textcolor{predicate}{\mathtt{r}}(\textcolor{variable}{\mathsf{X}})$ & $\top$ \\
        \hline
      \end{tabular}
    \end{column}
  \end{columns}
\end{frame}

% TODO: and how is a predicate with its arguments represented?
\begin{frame}{Variable Symmetry Breaking}
  \begin{block}{The Problem}
    Let $\{ \textcolor{variable}{\mathsf{W}}, \textcolor{variable}{\mathsf{X}},
    \textcolor{variable}{\mathsf{Y}} \}$ be the set of variables. Then
    \[
      \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}})
      \ifff
      \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{Y}}),
      \textcolor{predicate}{\mathtt{influences}}(\textcolor{variable}{\mathsf{Y}},
      \textcolor{variable}{\mathsf{X}})
    \]
    is equivalent to
    \[
      \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{Y}})
      \ifff
      \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}}),
      \textcolor{predicate}{\mathtt{influences}}(\textcolor{variable}{\mathsf{X}},
      \textcolor{variable}{\mathsf{Y}})
    \]
    and to
    \[
      \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{W}})
      \ifff
      \textcolor{predicate}{\mathtt{smokes}}(\textcolor{variable}{\mathsf{X}}),
      \textcolor{predicate}{\mathtt{influences}}(\textcolor{variable}{\mathsf{X}},
      \textcolor{variable}{\mathsf{W}})
    \]
  \end{block}
\end{frame}
